---
title: "Experiment 3: Mask-and-impute on Ropers2021 and Habowski2020"
output: html_notebook
---

## Impute Ropers2021 and Habowski2020 with Pirat, Pirat-T and PIrat-S

Run script **experiment3.R** to impute Ropers2021 and Habowski2020 datasets with Pirat-T and Pirat-S. Pseudo MVs, either MNAR or MCAR, are first generated with different seeds on the original datasets. Then, the resulting datasets containing pseudo MVs are imputed with these methods and stored in **res/**.

## Required packages

```{r}
library(ggplot2)
library(latex2exp)
```

## Compute errors of different methods

```{r}
# data.name = c("MouseColon", "RESET")
# impmethods = c("MLEMNAR_full", "MLEMNAR_TPG1", "MLEMNAR_PG")
# path2exp = file.path("../../2020-proteomics-transcriptomics/experiments/")
data.name = c("Habowski2020", "Ropers2021")
impmethods = c("Pirat", "Pirat-S", "Pirat-T")
path2exp = file.path("res")
mnar.perc = c(0, 1)
settingmv = paste("MNAR", mnar.perc)
seeds = as.character(0:9 + 543210)
methodfiles = paste(impmethods, ".rds", sep = "")
n_meth = length(impmethods)

rmses = data.frame(list())


for (cur.data.name in data.name) {
  dataset = file.path(path2exp, cur.data.name)
  for (settingmv.cur in settingmv) {
    for (seed in seeds) {
      print(seed)
      path2data = file.path(dataset, settingmv.cur, seed, "DATA.rds")
      gtdata = readRDS(path2data)
      n_cc = ncol(gtdata$adj)
      rmse_temp2 = data.frame(list())
      for (method in impmethods) {
        rmses_temp = data.frame(list())
        path2res = file.path(dataset, settingmv.cur,
                             seed, paste(method, ".rds", sep=''))
        resimp = readRDS(path2res)
        for (icc in 1:n_cc) {
          idx_pep = which(gtdata$adj[, icc] == 1)
          cc_size = length(idx_pep)
          Xcomp = gtdata$comp_pep_abs[, idx_pep]
          Xpseudo = gtdata$peptides_ab[, idx_pep]
          Ximp = resimp[, idx_pep]
          idx_pseudo_mis = which(!(is.na(Xcomp)) & is.na(Xpseudo), arr.ind = T)
          if (length(idx_pseudo_mis) != 0) {
            imp_vals = Ximp[idx_pseudo_mis]
            gt_vals = Xcomp[idx_pseudo_mis]
            abs_errs = abs(imp_vals - gt_vals)
            npseudo_val = length(imp_vals)
            frame2concat = data.frame(list(dataset = cur.data.name,
                                           mnar_perc = settingmv.cur,
                                           method = rep(method, npseudo_val),
                                           seed = rep(seed, npseudo_val),
                                           cc_size = rep(cc_size, npseudo_val), 
                                           gt_val = gt_vals, imp_val = imp_vals, 
                                           abs_err = abs_errs))
            rmses_temp = rbind(rmses_temp, frame2concat)
          }
        }
        rmse_temp2 = rbind(rmse_temp2, rmses_temp)
      }
      rmses = rbind(rmses, rmse_temp2)
    }
  }
}


rmses$cc_size = as.factor(rmses$cc_size)
rmses = rmses[order(rmses$cc_size), ]
# rmses$method[rmses$method == "MLEMNAR_full"] = "Pirat"
# rmses$method[rmses$method == "MLEMNAR_PG"] = "Pirat-T"
# rmses$method[rmses$method == "MLEMNAR_TPG1"] = "Pirat-S"
rmses$mnar_perc[rmses$mnar_perc == "MNAR 0"] = "MCAR"
rmses$mnar_perc[rmses$mnar_perc == "MNAR 1"] = "MNAR"
# rmses$dataset[rmses$dataset == "RESET"] = "Ropers2021"
# rmses$dataset[rmses$dataset == "MouseColon"] = "Habowski2020"
```

## Boxplots absolute error minus median

```{r}
rmses_trunc = rmses[rmses$cc_size %in% c(1) & rmses$method == "Pirat", ] # & rmses$method %in% c("Ours_MNAR", "Ours_MNAR_PG", "ImpSeq") , ] # & !(rmses$method %in% c("CenseMLE_cov_ratio_02")),]
# rmses_all = rmses
# rmses_all$cc_size = rmses_all$cc_size != 1
errors.not.pg1 = rmses$abs_err[rmses$dataset == "Ropers2021" & rmses$method == "Pirat" & rmses$cc_size != 1 & rmses$mnar_perc == "MCAR"]
abs_errs = rmses_trunc$abs_err[rmses_trunc$dataset == "Ropers2021" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MCAR"]
rmses_trunc$abs_err[rmses_trunc$dataset == "Ropers2021" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MCAR"] = abs_errs - median(errors.not.pg1)

errors.not.pg1 = median(rmses$abs_err[rmses$dataset == "Ropers2021" & rmses$method == "Pirat" & rmses$cc_size != 1 & rmses$mnar_perc == "MNAR"])
abs_errs = rmses_trunc$abs_err[rmses_trunc$dataset == "Ropers2021" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MNAR"]
rmses_trunc$abs_err[rmses_trunc$dataset == "Ropers2021" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MNAR"] = abs_errs - median(errors.not.pg1)

med.errors.not.pg1 = median(rmses$abs_err[rmses$dataset == "Habowski2020" & rmses$method == "Pirat" & rmses$cc_size != 1 & rmses$mnar_perc == "MCAR"])
abs_errs = rmses_trunc$abs_err[rmses_trunc$dataset == "Habowski2020" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MCAR"]
rmses_trunc$abs_err[rmses_trunc$dataset == "Habowski2020" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MCAR"] = abs_errs - med.errors.not.pg1

med.errors.not.pg1 = median(rmses$abs_err[rmses$dataset == "Habowski2020" & rmses$method == "Pirat" & rmses$cc_size != 1 & rmses$mnar_perc == "MNAR"])
abs_errs = rmses_trunc$abs_err[rmses_trunc$dataset == "Habowski2020" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MNAR"]
rmses_trunc$abs_err[rmses_trunc$dataset == "Habowski2020" & rmses_trunc$method == "Pirat" & rmses_trunc$cc_size == 1 & rmses_trunc$mnar_perc == "MNAR"] = abs_errs - med.errors.not.pg1

# rmses_trunc$mnar_perc[rmses_trunc$mnar_perc == "MNAR 0"] = "MCAR"
# rmses_trunc$mnar_perc[rmses_trunc$mnar_perc == "MNAR 1"] = "MNAR"
# 
# rmses_trunc$dataset[rmses_trunc$dataset == "RESET"] = "Ropers2021"
# rmses_trunc$dataset[rmses_trunc$dataset == "MouseColon"] = "Habowski2020"


p <- ggplot(rmses_trunc, aes(x=dataset, y=abs_err, fill=mnar_perc)) + 
  geom_boxplot(show.legend = TRUE)
p <- p + ylab(TeX("$\\Delta_{AE}$"))
p <- p + geom_hline(yintercept=0, linetype="dashed")
p <- p + coord_cartesian(ylim = c(min(rmses_trunc$abs_err), 2))
# p <- p + ggtitle(paste("Standardized Absolute Errors"))
p <- p + theme(panel.background = element_blank(),
               axis.title.x = element_blank())
p <- p + guides(fill=guide_legend(title="Scenario"))
p
```
## Boxplot Pirat absolute errors on singleton/non-singleton PGs 

```{r}
rmses_trunc = rmses[rmses$dataset == "Habowski2020" & rmses$method == "Pirat", ]
rmses_trunc$cc_size = as.integer(levels(rmses_trunc$cc_size)[rmses_trunc$cc_size])
rmses_trunc$cc_size = factor(rmses_trunc$cc_size > 1, labels = c("1", "> 1"))
p <- ggplot(rmses_trunc, aes(x=mnar_perc, y=abs_err, fill=cc_size)) + 
  geom_boxplot(show.legend = TRUE)
p <- p + ylab("Absolute Errors")
# p <- p + coord_cartesian(ylim = c(min(rmses_trunc$abs_err), 2))
p <- p + ggtitle(paste("A - Habowski2020"))
p <- p + theme(panel.background = element_blank(),
               axis.title.x = element_blank())
p <- p + guides(fill=guide_legend(title="PG size"))
p
rmses_trunc = rmses[rmses$dataset == "Ropers2021" & rmses$method == "Pirat", ]
rmses_trunc$cc_size = as.integer(levels(rmses_trunc$cc_size)[rmses_trunc$cc_size])
rmses_trunc$cc_size = factor(rmses_trunc$cc_size > 1, labels = c("1", "> 1"))
p <- ggplot(rmses_trunc, aes(x=mnar_perc, y=abs_err, fill=cc_size)) + 
  geom_boxplot(show.legend = TRUE)
p <- p + ylab("Absolute Errors")
# p <- p + coord_cartesian(ylim = c(min(rmses_trunc$abs_err), 2))
p <- p + ggtitle(paste("B - Ropers2021"))
p <- p + theme(panel.background = element_blank(),
               axis.title.x = element_blank())
p <- p + guides(fill=guide_legend(title="PG size"))
p
```

## Boxplot absolute errors Pirat, Pirat-S and Pirat-T

```{r}
cols = c("Pirat-S" = "#F8766D",
         "Pirat-T" = "#00BA38",
         "Pirat" = "#619CFF")
rmses_trunc = rmses[rmses$cc_size %in% c(1) & rmses$dataset == "Habowski2020", ]
p <- ggplot(rmses_trunc, aes(x=mnar_perc, y=abs_err, fill=method)) + 
  geom_boxplot(show.legend = TRUE)
p <- p + scale_fill_manual(
  name = "Methods",
  values = cols
)
p <- p + ylab("Absolute errors")
# p <- p + coord_cartesian(ylim = c(0, 2))
p <- p + ggtitle(paste("A - Habowski2020"))
p <- p + theme(panel.background = element_blank(),
               axis.title.x = element_blank())
p <- p + guides(fill=guide_legend(title="Method"))
p

rmses_trunc = rmses[rmses$cc_size %in% c(1) & rmses$dataset == "Ropers2021", ]
p <- ggplot(rmses_trunc, aes(x=mnar_perc, y=abs_err, fill=method)) + 
  geom_boxplot(show.legend = TRUE)
p <- p + scale_fill_manual(
  name = "Methods",
  values = cols
)
p <- p + ylab("Absolute errors")
# p <- p + coord_cartesian(ylim = c(0, 2))
p <- p + ggtitle(paste("B - Ropers2021"))
p <- p + theme(panel.background = element_blank(),
               axis.title.x = element_blank())
p <- p + guides(fill=guide_legend(title="Method"))
p
```

## Histogram of number of MVs on peptides containing pseudo-MVs

```{r}
##############################
# Non-singleton PGs
##############################

########### MouseCOlon 0 MNAR

data.name = c("Habowski2020")
mnar.perc = c(0)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)
seeds = c("543210", "543211", "543212", "543213", "543214", "543215", "543216", "543217", "543218", "543219") 

pseudo.mv.type = c()
gt.data = readRDS("../data/habowski_comp_pg.rds")
for (seed in seeds) {
  print(seed)
  # path2data = file.path("../../2020-proteomics-transcriptomics/", "experiments", data.name, "pseudo_na_percent_1_mis", settingmv, seed, "DATA.rds")
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")
  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) > 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("A - Habowski2020", "MCAR", "(PG size > 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGgt1.pdf")),
       width = 5, height = 5,
       device = "pdf")

########### MouseCOlon 1 MNAR

mnar.perc = c(1)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)

pseudo.mv.type = c()
for (seed in seeds) {
  print(seed)
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")
  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.data = readRDS("../processed_data/all_data_comp_pg_MouseColon.rds")
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) > 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
table(pseudo.mv.type)
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("B - Habowski2020", "MNAR", "(PG size > 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGgt1.pdf")),
       width = 5, height = 5,
       device = "pdf")

######### RESET 0 MNAR

data.name = c("Ropers2021")
mnar.perc = c(0)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)
pseudo.mv.type = c()
gt.data = readRDS("../data/ropers_comp_pg.rds")
for (seed in seeds) {
  print(seed)
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")
  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) > 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
table(pseudo.mv.type)
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("C - Ropers2021", "MCAR", "(PG size > 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGgt1.pdf")),
       width = 5, height = 5,
       device = "pdf")

######### RESET 1 MNAR

mnar.perc = c(1)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)

pseudo.mv.type = c()
for (seed in seeds) {
  print(seed)
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")
  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) > 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
table(pseudo.mv.type)
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("D - Ropers2021", "MNAR", "(PG size > 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGgt1.pdf")),
       width = 5, height = 5,
       device = "pdf")

###############################
# Singleton PGs
###############################

########### MouseCOlon 0 MNAR


data.name = c("Haboswki2020")
mnar.perc = c(0)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)

pseudo.mv.type = c()
gt.data = readRDS("../data/habowski_comp_pg.rds")
for (seed in seeds) {
  print(seed)
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")
  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) == 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
table(pseudo.mv.type)
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("E - Habowski2020", "MCAR", "(PG size = 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGeq1.pdf")),
       width = 5, height = 5,
       device = "pdf")

########### MouseCOlon 1 MNAR

mnar.perc = c(1)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)

pseudo.mv.type = c()
for (seed in seeds) {
  print(seed)
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) == 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
table(pseudo.mv.type)
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("F - Habowski2020", "MNAR", "(PG size = 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGeq1.pdf")),
       width = 5, height = 5,
       device = "pdf")

######### RESET 0 MNAR

data.name = c("Ropers2021")
mnar.perc = c(0)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)
pseudo.mv.type = c()
gt.data = readRDS("../data/ropers_comp_pg.rds")
for (seed in seeds) {
  print(seed)
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")
  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) == 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
table(pseudo.mv.type)
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("G - Ropers2021", "MCAR", "(PG size = 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGeq1.pdf")),
       width = 5, height = 5,
       device = "pdf")

######### RESET 1 MNAR

mnar.perc = c(1)
settingmv = paste("MNAR", mnar.perc)
print(data.name)
print(settingmv)

pseudo.mv.type = c()
for (seed in seeds) {
  print(seed)
  path2data = file.path("res", data.name, settingmv, seed, "DATA.rds")
  pseudodata = readRDS(path2data)
  pseudo.ab = pseudodata$peptides_ab
  gt.ab = gt.data$peptides_ab 
  idx.pgs.1 = which(colSums(pseudodata$adj) == 1)
  idx.pep.pg.1 = which(rowSums(pseudodata$adj[, idx.pgs.1]) >= 1)
  pseudo.ab = pseudo.ab[, idx.pep.pg.1]
  idx.same.pep = which(colnames(gt.ab) %in% colnames(pseudo.ab))
  gt.ab = gt.ab[, idx.same.pep]
  mask.pseudo = is.na(pseudo.ab) & !is.na(gt.ab)
  n.mvs.mat = matrix(rep(colSums(is.na(pseudo.ab)), each=18), 18)
  pseudo.mv.type = c(pseudo.mv.type, n.mvs.mat[mask.pseudo])
}
table(pseudo.mv.type)
p <- ggplot(data = data.frame(N_MVs_per_pep = pseudo.mv.type), aes(x = N_MVs_per_pep)) +
  geom_bar() + ggtitle(paste("H - Ropers2021", "MNAR", "(PG size = 1)")) + 
  xlab("Number of MVs per peptide of each pseudo-MV")
p
ggsave(file.path("..", "experiments", "hist_N_MVs", paste0(data.name, settingmv, "PGeq1.pdf")),
       width = 5, height = 5,
       device = "pdf")

```

## Estimation of missingness mechanism

```{r}
# data.pep.rna.mis = readRDS("../../2020-proteomics-transcriptomics/processed_data/all_data_comp_pg_RESET.rds")

data.pep.rna.mis = readRDS("../data/ropers_comp_pg.rds")
mv_rates = colMeans(is.na(data.pep.rna.mis$peptides_ab))
mean_abund = colMeans(data.pep.rna.mis$peptides_ab, na.rm = T)
mean_abund_sorted = sort(mean_abund, index.return = T)
mv_rates_sorted = mv_rates[mean_abund_sorted$ix]
kernel_size = 10
probs = rep(0, length(mean_abund) - kernel_size + 1)
for (i in 1:length(probs)) {
  probs[i] = mean(mv_rates_sorted[i:(i + kernel_size - 1)])
}
not0 = probs != 0
m_ab_sorted = mean_abund_sorted$x[not0]
probs = probs[not0]
plot(m_ab_sorted[1:length(probs)], log(probs),
     main=paste0("Ropers2021, k=",kernel_size), ylab="log(p_mis)", 
     xlab="observed mean")
res.reg = lm(log(probs) ~ m_ab_sorted[1:length(probs)])
sum.reg.reg = summary(res.reg)
print(sum.reg.reg)
abline(res.reg, col="red")
mylabel = bquote(italic(R)^2 == .(format(summary(res.reg)$r.squared, digits = 3)))
text(x = 16, y = -2.5, labels = mylabel)

phi0 = -res.reg$coef[1]
phi = -res.reg$coef[2]
if ((phi <= 1e-3) | (sum.reg.reg$coefficients[2,4] > 0.001)) {
  phi = 0
}
print(paste("Gamma0 estimated = ", phi0))
print(paste("Gamma1 estimated = ", phi))
```
