---
title: "Experiment 2: Mask-and-impute on Capizzi2022 and Vilallongue2022"
output: html_notebook
---

## Impute Capizzi2020 and Vilallongue2020 datasets with various methods

Run script **experiment2.R** to impute Capizzi2022 and Vilallongue2022 datasets with SeqKNN, BPCA, ImpSeq, MinProb, QRILC, Pirat and Pirat_degenerated. Pseudo MVs with various MNAR ratio are first generated with different seeds on the complete dataset. Then, the resulting datasets containing pseudo MVs are imputed with these methods and stored in **res/**.

## Required packages

```{r}
library(ggplot2)
library(Pirat)
```


## Choose dataset of interest

### Capizzi2020

```{r}
name.data = "Capizzi2022"
path2res = file.path("res", name.data)
# path2res = file.path("..", "..", "2020-proteomics-transcriptomics", "experiments", name.data, "pseudo_na_percent_16")
name.title = "peptide-level, DDA, n=10"
```

### Vilallongue2020 SCN

```{r}
name.data = "Vilallongue2022_SCN"
path2res = file.path("res", name.data)
name.title = "precursor-level, DDA, n=8"
```


### Vilallongue2020 SC

```{r}
name.data = "Vilallongue2022_SC"
path2res = file.path("res", name.data)
name.title = "precursor-level, DDA, n=8"
```

## Error metrics functions

```{r}
compute_rmse = function(Ximp, Xcomp, Xref) {
  na_gen = is.na(Xref) & !is.na(Xcomp) & !is.na(Ximp)
  rmseval = sqrt(mean((Xcomp[na_gen] - Ximp[na_gen])^2))
  return(rmseval)
}

compute_mae = function(Ximp, Xcomp, Xref) {
  na_gen = is.na(Xref) & !is.na(Xcomp) & !is.na(Ximp)
  rmseval = mean(abs(Xcomp[na_gen] - Ximp[na_gen]))
  return(rmseval)
}

nbypep_rmse = function(Ximp, Xcomp, Xref) {
  na_gen = is.na(Xref) & !is.na(Xcomp) & !is.na(Ximp)
  vars = apply(Xcomp, 2, var)
  Ximp[!na_gen] = NA
  Xcomp[!na_gen] = NA
  serrs = (Xcomp - Ximp)^2
  sum_serrs_normalised = t(t(serrs) / vars)
  return(sqrt(mean(sum_serrs_normalised, na.rm = T)))
}

rel_rmse = function(Ximp, Xcomp, Xref) {
  na_gen = is.na(Xref) & !is.na(Xcomp) & !is.na(Ximp)
  rmseval = sqrt(mean(((Xcomp[na_gen] - Ximp[na_gen]) / Xcomp[na_gen])^2))
  return(rmseval)
}
```


## Compute RMSE and MAEs for each MNAR ratio/seed/method and aggregate over seeds

```{r}
mnar.perc = seq(0, 1, 0.25)
mnar.perc.files = paste("MNAR", mnar.perc)
seeds = as.character(0:4 + 543210)
methods2comp = c("Pirat", "MinProb", "QRILC", "SeqKNN", "ImpSeq", "BPCA", "Pirat_degenerated")
methodfiles = paste(methods2comp, ".rds", sep = "")

n_mnar_perc = length(mnar.perc)
n_seeds = length(seeds)
n_meth = length(methods2comp)

rmses = data.frame(expand.grid(list(mnar_perc=mnar.perc,
                                    seed=seeds,
                                    method=methods2comp), stringsAsFactors = F))
rmses$rmse = NA
rmses$n_by_pep_rmse = NA
rmses$rel_rmse = NA
rmses$mae = NA
rmses$time_comp = NA
for (i in 1:length(mnar.perc)) {
  for (j in 1:length(seeds)) {
    pathdata = file.path(path2res, mnar.perc.files[i], seeds[j], "DATA.rds")
    print(pathdata)
    data.pep = readRDS(pathdata)
    idx.prot.not.singl = which(colSums(data.pep$adj) >= 1) # PG size
    idx.pep.not.singl = which(rowSums(data.pep$adj[, idx.prot.not.singl]) >= 1) # Dont touch this line
    complete_ab = data.pep$comp_pep_abs[, idx.pep.not.singl]
    pseudo_abs = data.pep$peptides_ab[, idx.pep.not.singl]
    
    for (k in 1:length(methodfiles)) {
      # print(methodfiles[k])
      pathresmeth = file.path(path2res, mnar.perc.files[i], seeds[j], methodfiles[k])
      if (methodfiles[k] %in% list.files(file.path(path2res, mnar.perc.files[i], seeds[j]))) {
        resimp = readRDS(pathresmeth)[, idx.pep.not.singl]
        # print(all(!is.na(resimp)))
        rmse = compute_rmse(resimp, complete_ab, pseudo_abs)
        idx_row = which(rmses$mnar_perc == mnar.perc[i] & rmses$seed == seeds[j] & rmses$method == methods2comp[k])
        rmses$rmse[idx_row] = rmse
        peprmse = nbypep_rmse(resimp, complete_ab, pseudo_abs)
        rmses$n_by_pep_rmse[idx_row] = peprmse
        relative_rmse = rel_rmse(resimp, complete_ab, pseudo_abs)
        rmses$rel_rmse[idx_row] = relative_rmse
        mae = compute_mae(resimp, complete_ab, pseudo_abs)
        rmses$mae[idx_row] = mae
        pathtimemeth = file.path(path2res, mnar.perc.files[i], seeds[j], paste(methods2comp[k], "_time.rds", sep=""))
        timecomp = readRDS(pathtimemeth)
        rmses$time_comp[idx_row] = timecomp
      }
    }
  }
}
mean_rmses = data.frame(list(mnar_perc = rep(mnar.perc, each=n_meth),
                             Method = rep(methods2comp, n_mnar_perc)))
mean_rmses$rmse = NA
mean_rmses$n_by_pep_rmse = NA
mean_rmses$rel_rmse = NA
mean_rmses$mae = NA
mean_rmses$sd_rmse = NA
mean_rmses$sd_n_by_pep_rmse = NA
mean_rmses$sd_rel_rmse = NA
mean_rmses$sd_mae = NA
col2replace = c("rmse", "n_by_pep_rmse", "rel_rmse", "mae")
sd_col2replace = c("sd_rmse", "sd_n_by_pep_rmse", "sd_rel_rmse", "sd_mae")
for (i in 1:length(mnar.perc)) {
  for (k in 1:length(methodfiles)) {
    idx_row = which(rmses$mnar_perc == mnar.perc[i] & rmses$method == methods2comp[k])
    idx2replace = which(mean_rmses$mnar_perc == mnar.perc[i] & mean_rmses$Method == methods2comp[k])
    mean_rmses[idx2replace, col2replace] = colMeans(rmses[idx_row, col2replace], na.rm = F)
    mean_rmses[idx2replace, sd_col2replace] = apply(rmses[idx_row, col2replace], 2, sd, na.rm = F)
  }
}
```

## Color for curves

```{r}
cols = c("Pirat"="blue", 
         "QRILC"="violet", 
         "BPCA"="red", 
         "ImpSeq" = "green", 
         "SeqKNN" = "yellow3",
         "msImpute_mnar" = "orange",
         "msImpute_mar" = "olivedrab",
         "MinProb" = "lightskyblue",
         "trKNN" = "mediumpurple",
         "GMS" = "darkmagenta",
         "ImpSeqRob" = "peachpuff3",
         "MissForest" = "slategray",
         "SVD" = "darkgreen",
         "LLS" = "saddlebrown",
         "MLE" = "darkturquoise",
         "KNN" = "lightcoral",
         "Pirat_degenerated" = "grey")
```


## Display MAE curves


```{r}
methods2disp = c("Pirat", "Pirat_degenerated", "SeqKNN", "ImpSeq", "BPCA") # Choose imputation methods to display
mean_rmses_best = mean_rmses[mean_rmses$Method %in% methods2disp,]
p <- ggplot(mean_rmses_best, aes(x=mnar_perc, y=mae, color=Method, fill = Method)) + 
   geom_point() + geom_line() +
  geom_ribbon(aes(ymin=mae-sd_mae, ymax=mae+sd_mae),
              linetype = 0, alpha=0.15)
p <- p + xlab("MNAR proportion") + ylab("Average MAE") + 
  ggtitle(paste(name.data, name.title)) +
  scale_color_manual(
    values = cols,
    breaks = methods2disp
  ) +
  scale_fill_manual(
    values = cols,
    breaks = methods2disp
  ) +
  theme(panel.background = element_blank(),
        legend.position = c(0.23, 0.75),
        legend.title.align=0.5,
        legend.title = element_text(size=11),
        legend.text = element_text(size=11),
        legend.key = element_rect(fill = "white"))
p
ggsave(file.path(path2res, paste0("mae_mnar", name.data, ".pdf")),
  width = 5, height = 5,
  device = "pdf") # To save figure
```

## Display RMSE curves

```{r}
methods2disp = c("Pirat", "Pirat_degenerated", "SeqKNN", "ImpSeq", "BPCA") # Choose imputation methods to display
mean_rmses_best = mean_rmses[mean_rmses$Method %in% methods2disp,]
p <- ggplot(mean_rmses_best, aes(x=mnar_perc, y=rmse, color=Method, fill = Method)) + 
  geom_point() + geom_line() +
  geom_ribbon(aes(ymin=rmse-sd_rmse, ymax=rmse+sd_rmse),
              linetype = 0, alpha=0.15)
p <- p + xlab("MNAR proportion") + ylab("Average RMSE") + 
  ggtitle(paste(name.data, name.title)) +
  scale_color_manual(
    values = cols,
    breaks = methods2disp
  ) +
  scale_fill_manual(
    values = cols,
    breaks = methods2disp
  ) +
  theme(panel.background = element_blank(),
        legend.position = c(0.75, 0.27),
        legend.title.align=0.5,
        legend.title = element_text(size=11),
        legend.text = element_text(size=11),
        legend.key = element_rect(fill = "white"))
p
ggsave(file.path(path2res, paste0("rmse_mnar", name.data, ".pdf")),
  width = 5, height = 5,
  device = "pdf") # To save figure
```
## Within-PG correlations

```{r}
data.name = "Capizzi2022"
pep.data.comp = readRDS("../data/capizzi_comp.rds")
# data.name = "Vilallongue2022_SCN"
# pep.data.comp = readRDS("../data/vilallongue_scn_ion_vsn.rds")
# data.name = "Vilallongue2022_SC"
# pep.data.comp = readRDS("../data/vilallongue_sc_ion_vsn.rds")
p = plot_pep_correlations(pep.data.comp, titlename = data.name)
```


