---
title: "Experiment 1: differential analysis on benchmark datasets"
output: html_notebook
---


## Impute benchmark datasets with various methods

Run script **experiment1.R** to impute Huang2020, Cox2014 and Bouyssié2020 benchmark datasets with 16 different imputation methods, including Pirat. Imputed datasets are stored in **res/**.

## Load required packages

```{r}
library(pROC)
library(ggplot2)
```


## Choose dataset to study

Execute the cell among the 3 following corresponding to the dataset of interest.

### Huang2020

```{r}
data.name = "Huang2020"
path2res = file.path("res", data.name)
pathdata = file.path("..", "data", "Huang2020_data_comp.rds")
title_roc = "A - Huang2020, precursor-level, DIA, n=25"
```

### Cox2014

```{r}
data.name = "Cox2014"
path2res = file.path("res", data.name)
pathdata = file.path("..", "data", "cox_comp.rds")
title_roc = "B - Cox2014, peptide-level, DDA, n=8"
```


### Bouyssié2020

```{r}
# data.name = "Proline"
# path2res = file.path("../../2020-proteomics-transcriptomics/experiments/", data.name, "543210_w_shared")
# pathdata = file.path(path2res, "DATA.rds")
data.name = "Bouyssie2020"
path2res = file.path("res", data.name)
pathdata = file.path("..", "data", "bouyssie_comp.rds")
title_roc = "C - Bouyssie2020, peptide-level, DDA, n=40"
```

## Compute p-values for each peptide and for each imputation method

```{r}
methods2comp = c("KNN", "GMS", "ImpSeqRob", "Pirat", "MissForest", "MinProb", "QRILC", "SVD", "LLS", "trKNN", "SeqKNN", "ImpSeq", "BPCA", "MLE", "msImpute_mar", "msImpute_mnar")
data.pep = readRDS(pathdata)
npeps = ncol(data.pep$peptides_ab)
rocstats = data.frame(list(diff.pep = factor(data.pep$mask_pep_diff)))
groups = as.factor(rep(1:10, each = 4))

for (resfile in methods2comp) {
  pathresmeth = file.path(path2res, paste(resfile, ".rds", sep = ""))
  print(pathresmeth)
  resimp = readRDS(pathresmeth)
  pvals = apply(resimp, 2, function(x) {
    pval = tryCatch(summary(aov(ab~group, data.frame(list(group = groups, ab = x))))[[1]][1,5],
                    error = function(cond) {print(cond); return(1.)});
    return(if (is.null(pval)) 1. else pval)})
  rocstats[resfile] = pvals
}

saveRDS(rocstats, file.path(path2res, "rocstats.RDS")) # better to save as computation is a bit long
# rocstats = readRDS(file.path(path2res, "rocstats.RDS"))

rocs = list()
for (i in 1:length(methods2comp)) {
  rocs[[methods2comp[i]]] = roc(rocstats$diff.pep, rocstats[, methods2comp[i]])
}
```

## Assessing performances

### Method color code

```{r}
cols = c("MinProb" = "lightskyblue",
         "GMS" = "darkmagenta",
         "ImpSeqRob" = "peachpuff3",
         "MissForest" = "slategray",
         "SVD" = "darkgreen",
         "LLS" = "saddlebrown",
         "MLE" = "darkturquoise",
         "KNN" = "lightcoral",
         "SeqKNN" = "yellow3",
         "QRILC"="violet",
         "BPCA"="red", 
         "ImpSeq" = "green", 
         "trKNN" = "mediumpurple",
         "msImpute_mar" = "olivedrab",
         "msImpute_mnar" = "orange",
         "Pirat" = "blue")
```


### ROC

```{r}
fplim = 0.99 # Limit threshold for false positive rate

g = NULL
g <- ggroc(rocs)
g <- g + ggtitle(title_roc) +
  scale_color_manual(
      name = "Methods",
      values = cols,
      breaks = sort(names(rocs))
  ) +
  theme(panel.background = element_blank(),
        aspect.ratio = 1) + 
  xlim(c(1, fplim)) 
g
```

###  Precision Recall (PR) Curve

```{r}
get.prec.recall = function(roc.cur) {
  roc.all = coords(roc.cur, ret = "all", transpose = FALSE)
  prec.rec = roc.all[, c("recall", "precision")]
  return(prec.rec)
}
prs = lapply(rocs, get.prec.recall)

concat.prs = do.call(rbind, prs)
concat.prs$Method = factor(rep(names(prs), times = unlist(lapply(prs, nrow))),
                           levels = names(cols))

g <- ggplot(data = concat.prs, aes(x = recall, y = precision, color = Method)) +
  geom_line()
g <- g + ggtitle(title_roc) +
  scale_color_manual(
    values = cols,
  ) + ylab("Precision (1 - FDP)") + xlab("Recall (TPR)") +
  theme(panel.background = element_blank(),
        legend.position = "none",
        # legend.position = c(0.8, 0.3), legend.key.width = unit(1, "cm"), 
        # legend.text = element_text(size=11), 
        # legend.title = element_text(size=11),
        # legend.title.align=0.5,
        legend.key = element_rect(fill = "white"),
        aspect.ratio = 1)
g <- g + ylim(c(0.9, 1))
g <- g + geom_hline(yintercept=0.95, linetype = 3) + geom_hline(yintercept=0.99, linetype = 3)
ggsave(file.path(path2res, paste0("pr_curve_", data.name, ".pdf")),
  width = 5, height = 5,
  device = "pdf") # To save figure
```

### AUCPR

```{r}
compute.aucpr = function(pr.frame) {
  pr = pr.frame$precision
  pr[is.na(pr)] = 1
  rec = pr.frame$recall
  aucpr = sum(pr[-length(pr)] * (rec[-length(rec)] - rec[-1]))
  return(aucpr)
}

aucprs.c = lapply(prs, compute.aucpr)
aucprs.c = unlist(aucprs.c)
aucprs.c
```